### 2013-03-25 doc/dissertation/document/chapters/mc_background/mc_background.tex
adding adjoint evolution of a solution to mc background

As a means of visually demonstrating the direct Monte Carlo method,
consider a 2-dimensional thermal diffusion problem with sources on the
left and right hand sides of the domain and a smaller uniform source
as shown in Figure~\ref{fig:heat_setup}.
\begin{figure}[h!]
  \begin{center}
    \scalebox{1.0}{ \input{heat_eq_setup.pdftex_t} }
  \end{center}
  \caption{\textbf{Problem setup for 2D heat equation.}
    \textit{Dirichlet conditions are set for the temperature on all 4
      boundaries of the Cartesian grid. Background source of 1.0
      present. $50 \times 50$ grid.}}
  \label{fig:heat_setup}
\end{figure}
For this problem, the number of histories used to compute the solution
at each state in the domain was increased from 1 to 1000 in order to
show its effects on the solution and the statistical nature of the
method. Figure~\ref{fig:direct_evolution} gives these results.
\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=4in]{direct_evolution.png}
  \end{center}
  \caption{\textbf{Direct Monte Carlo solution to the heat equation
      with varying numbers of histories.} \textit{Top left: 1 history
      per state. Top right: 10 histories per state. Bottom left: 100
      histories per state. Bottom right: 1000 histories per state.}}
  \label{fig:direct_evolution}
\end{figure}
As the number of histories used per state is increased, the
statistical variance of the solutions is decreased as more tallies are
made. At 1000 histories per state, enough tallies have been made to
generate a reasonable estimate for the structure of the solution.

@@

>>>>

As a means of visually demonstrating the direct Monte Carlo method,
consider a 2-dimensional thermal diffusion problem with sources on the
left and right hand sides of the domain and a smaller uniform source
as shown in Figure~\ref{fig:heat_setup}.
\begin{figure}[h!]
  \begin{center}
    \scalebox{1.0}{ \input{chapters/mc_background/heat_eq_setup.pdftex_t} }
  \end{center}
  \caption{\textbf{Problem setup for 2D heat equation.}
    \textit{Dirichlet conditions are set for the temperature on all 4
      boundaries of the Cartesian grid. Background source of 1/5 the
      value of the boundary sources present. $50 \times 50$ grid.}}
  \label{fig:heat_setup}
\end{figure}
For this problem, the number of histories used to compute the solution
at each state in the domain was increased from 1 to 1000 in order to
show its effects on the solution and the statistical nature of the
method. Figure~\ref{fig:direct_evolution} gives these results.
\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=5in]{chapters/mc_background/direct_evolution.png}
  \end{center}
  \caption{\textbf{Direct Monte Carlo solution to the heat equation
      with varying numbers of histories.} \textit{Top left: 1 history
      per state. Top right: 10 histories per state. Bottom left: 100
      histories per state. Bottom right: 1000 histories per state.}}
  \label{fig:direct_evolution}
\end{figure}
As the number of histories used per state is increased, the
statistical variance of the solutions is decreased as more tallies are
made. At 1000 histories per state, enough tallies have been made to
generate a reasonable estimate for the structure of the solution.

@@

As a means of visually demonstrating the adjoint Monte Carlo method,
again considerthe a 2-dimensional thermal diffusion problem with
sources on the left and right hand sides of the domain and a smaller
uniform source as shown in Figure~\ref{fig:heat_setup}. For the
adjoint method, the number of histories sampled from the source was
increased from 10 to 10,000,000 in order to show its effects on the
solution and the statistical nature of the
method. Figure~\ref{fig:adjoint_evolution} gives these results.
\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=5in]{chapters/mc_background/adjoint_evolution.png}
  \end{center}
  \caption{\textbf{Adjoint Monte Carlo solution to the heat equation
      with varying numbers of histories.} \textit{Top left: 10
      histories per state. Top right: 1,000 histories per
      state. Bottom left: 100,000 histories per state. Bottom right:
      10,000,000 histories per state.}}
  \label{fig:adjoint_evolution}
\end{figure}
As the number of histories used per state is increased, the
statistical variance of the solutions is decreased as more tallies are
made. At 10,000,000 histories per state, enough tallies have been made
to generate a reasonable estimate for the structure of the
solution. The visual difference between
Figures~\ref{fig:direct_evolution} and \ref{fig:adjoint_evolution} is
precisely that determined by their mathematics. As the adjoint
solution evolves with the addition of histories, more histories
emanate from the source with more penetrating the domain and making
contributions to those tallies. As the direct method evolves, the
solution simply comes into focus as histories are emanated from each
state in the system and contribute only to the solution in that state
regardless of their path.
